### 2020-08-28 01_background.Rmd
remove functions

---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Getting started

This section covers:

1. Installing the `R` package `atlastools`,

2. Simulating some realistic looking movement data using the `R` package `smoove` from [@gurarie2017], and

3. Introducing positioning errors into the simulated movement data.

## Installing `atlastools`

This paper refers extensively to the `R` package `atlastools` (Gupte 2020), which can be installed from [Github.](https://github.com/pratikunterwegs/atlastools). 
See the repository for information on the build status.
The code chunk below shows how to install `atlastools`.

```{r install_package, eval=FALSE}
# use either devtools or remotes to install
install.packages("devtools")

# installation using devtools
devtools::install_github("pratikunterwegs/atlastools")
```

## Simulating movement data

Here, we simulate some movement data using the `smoove` `R` package from [@gurarie2017].

```{r prep_libs_01_02}
# load smoove and datatable
library(smoove)
library(data.table)
```

```{r save_data}
data <- do_smoove_data()

# save simulated data
fwrite(data, "data/data_sim.csv")
```

Here, we write functions to add normally distributed errors to the position coordinates of the simulated data.

Example of introduced error.

```{r}
# load data
data <- fread("data/data_sim.csv")

# introduce errors to data
data[, `:=`(x_error = do_add_error(x, std_dev = 0.01),
            y_error = do_add_error(y, std_dev = 0.01))]

# plot real and introduced error
ggplot()+
  geom_path(data = data, aes(x, y),
            col = "blue")

atlastools::atl_compare_tracks(d2, data)
```