### 2020-10-05 11_residence_patch.Rmd
some code edits

---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Residence patches and their construction

## Prepare libraries {-}

```{r prep_libs_02_01}
library(data.table)
library(atlastools)
library(ggplot2)

# prepare a palette
pal <- RColorBrewer::brewer.pal(4, "Set1")
```

## An example with simulated data

### Read data and classify

```{r}
# read in the data
data <- fread("data/data_sim.csv")[5000:10000, ]
data[, id := "a"]

# make residence patch
patch <- atl_res_patch(data, 
                       buffer_radius = 0.0001, 
                       lim_spat_indep = 0.02, 
                       lim_time_indep = 3)

# get spatial representation
patch_sf <- atl_patch_summary(patch_data = patch, 
                              which_data = "spatial", 
                              buffer_radius = 0.01)

# get summary data
patch_summary <- atl_patch_summary(patch_data = patch,
                                   which_data = "summary")
```

### Plot classified residence patches

```{r}
plot_patches <-
  ggplot()+
  geom_sf(data = patch_sf,
          colour = "grey20",
          fill = "grey90",
          show.legend = FALSE)+
  geom_point(data = patch_summary,
             aes(x_median + 0.5, 
                 y_median,
                 size = duration),
             # size = 8,
             shape = 2,
             col = pal[2],
             stroke = 2,
             show.legend = F)+
  geom_path(data = data,
             aes(x, y),
            size = 0.3)+
  geom_curve(data = patch_summary,
            aes(x = x_end + 0.5, y = y_end,
                xend = shift(patch_summary$x_start, 
                             type = "lag") + 0.5,
                yend = shift(patch_summary$y_start, 
                             type = "lag")),
            col = "grey30",
            size = 0.3,
            ncp = 2,
            curvature = -0.4,
            arrow = arrow(angle = 20,
                          length = unit(2, units = "mm"), 
                          type = "closed", ends = "first"))+
  annotate(geom = "text",
           x = c(0.5, 1.1),
           y = 1,
           label = sprintf("(%s)", c("a", "b")),
           fontface = "bold")+
  scale_colour_distiller(palette = "YlGnBu",
                         trans = "log10")+
  coord_sf()+
  ggthemes::theme_map()+
  theme(legend.position = "right")

plot_patches

# save figure
ggsave(plot_patches, 
       filename = "figures/fig_residence_patch.png",
       width = 170, height = 100, units = "mm")
```