### 2020-10-28 06_bat_data.Rmd
add prelim bat data code

---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Processing calibration data

## Prepare libraries

```{r}
# load libs
library(data.table)
library(RSQLite)
library(atlastools)
library(ggplot2)
library(patchwork)

# prepare a palette
pal <- RColorBrewer::brewer.pal(4, "Set1")
```

## Read bat data

Read the bat data and convert to a csv file.

```{r read_bat_data}
# prepare the connection
con <- dbConnect(drv = SQLite(), dbname = "data/Three_example_bats.sql")

# list the tables
table_name <- dbListTables(con)

# prepare to query all tables
query <- sprintf('select * from \"%s\"', table_name)

# query the database
data <- dbGetQuery(conn = con, statement = query)

# disconnect from database
dbDisconnect(con)
```

Convert data to csv.

```{r}
# convert data to datatable
setDT(data)

# write data for QGIS
fwrite(data, file = "data/bat_data.csv")
```

## WIP