### 2020-10-28 01_background.Rmd

---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Getting started

This section covers:

1. Installing the `R` package `atlastools`,

2. Simulating some realistic looking movement data using the `R` package `smoove` from [@gurarie2017], and

3. Introducing positioning errors into the simulated movement data.

## Installing `atlastools`

This paper refers extensively to the `R` package `atlastools` (Gupte 2020), which can be installed from [Github](https://github.com/pratikunterwegs/atlastools).

Releases of the package from Github can be found on Zenodo: (https://doi.org/10.5281/zenodo.4033155)

The code chunk below shows how to install `atlastools`.

```{r install_package, eval=FALSE}
# use either devtools or remotes to install
install.packages("devtools")

# installation using devtools
devtools::install_github("pratikunterwegs/atlastools")
```

## Simulating movement data

Here, we simulate some movement data using the `smoove` `R` package from [@gurarie2017].

First, we load `smoove` and `data.table`, as well as some helper functions that make use of them to simulate data for use.

```{r prep_libs_01_02, eval=FALSE}
# load smoove and datatable
library(smoove)
library(data.table)

# source helper functions
source("R/helper_functions.R")
```

```{r save_data}
data <- do_smoove_data()
# save simulated data
fwrite(data, "data/data_sim.csv")
```

## Simualte for residence patches

```{r}
# do smoove data using a RACVM
data <- smoove::simulateRACVM(dt = 0.1, Tmax = 500, omega = 5, v0 = 1, mu = 0.1)
data <- as.data.table(data$XY)

# assign id
data[, id := "test"]
data[, time := seq_len(nrow(data))]
# rename x and y
setnames(data, old = c("x", "y"), new = c("y", "x"))

# save data
fwrite(data, "data/data_for_res_patch.csv")
```