### 2020-11-16 index.rmd
update index

--- 
knit: "bookdown::render_book"
title: "Supplementary Material for: A Rough Guide to Pre-processing High-Frequency Animal Tracking Data"
author:
    - Pratik R. Gupte
    - Christine E. Beardsworth
    - Orr Spiegel
    - Emmanuel Lourie
    - Allert Bijleveld
date: "`r Sys.Date()`"
output: bookdown::pdf_document2
documentclass: article
toc-depth: 2
secnumdepth: 2
geometry: "left=4cm, right=3cm, top=2.5cm, bottom=2.5cm"
bibliography: [references.bib]
link-citations: yes
github-repo: pratikunterwegs/atlas-best-practices
description: "Supplementary Material for: A Rough Guide to Pre-processing High-Frequency Animal Tracking Data"
---