### 2020-11-22 _05_calibration_data.Rmd
re add res time

---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Processing calibration data

## Prepare libraries

```{r}
# load libs
library(data.table)
library(atlastools)
library(ggplot2)
library(patchwork)

# prepare a palette
pal <- RColorBrewer::brewer.pal(4, "Set1")
```

## Preliminary visualisation

```{r}
# read and plot example data
data <- fread("data/atlas1060_allTrials_annotated.csv")
data_raw <- copy(data)
```

```{r echo=FALSE}
# plot data
fig_data_raw <-
  ggplot(data)+
  geom_path(aes(x, y),
             col = "grey", alpha = 0.5, size = 0.3)+
  geom_point(aes(x, y),
             col = pal[3], alpha = 0.2, size = 2)+
  ggthemes::theme_map()+
  coord_sf(crs = 32631)

# save figure
ggsave(fig_data_raw, filename = "figures/fig_calibration_raw.png",
       width = 185 / 25)
```

## Filter by bounding box

Save an unprocessed copy.

```{r}
data_unproc <- copy(data)
```

Filter by a bounding box.

```{r}
# remove inside must be set to falses
data <- atl_filter_bounds(data = data, 
                          x = "x", y = "y", 
                          x_range = c(645000, max(data$x)), 
                          remove_inside = FALSE)
```

```{r echo=FALSE}
# plot data
fig_data_bbox <-
  ggplot()+
  geom_path(data = data_raw,
            aes(x, y),
             col = "grey", alpha = 0.5, size = 0.3)+
  geom_point(data = data_raw,
             aes(x, y,
                 col = x > 645000),
             alpha = ifelse(data_raw$x > 645000, 0.5, 1),
             size = ifelse(data_raw$x > 645000, 0.3, 2),
             show.legend = F)+
  scale_colour_manual(values = c("grey", pal[3]))+
  geom_vline(xintercept = 645000,
             col = "grey", lty = 2)+
  ggspatial::annotation_scale(location = "br")+
  ggthemes::theme_few()+
  theme(axis.text = element_blank(),
        axis.title = element_blank())+
  coord_sf(crs = 32631)

# save result
ggsave(fig_data_bbox, filename = "figures/fig_calib_bbox.png",
       width = 185 / 25)
```

## Filter trajectories

### Handle time

Time in ATLAS tracking is counted in milliseconds and is represented by a 64-bit integer (type `long`), which is not natively supported in R; it will instead be converted to a `numeric`, or `double`.

This is not what is intended, but it works. The `bit64` package can help handle 64-bit integers if you want to keep to intended type.

A further issue is that 64-bit integers (whether represented as `bit64` or `double`) do not yield meaninful results when you try to convert them to a date-time object, such as of the class `POSIXct`.

This is because `as.POSIXct` fails when trying to work with 64-bit integers (it cannot interpret this type), and returns a date many thousands of years in the future (approx. 52,000 CE) if the time column is converted to `numeric`.

There are two possible solutions. The parsimonious one is to convert the 64-bit number to a 32-bit short integer (dividing by 1000), or to use the `nanotime` package.

The conversion method loses an imperceptible amount of precision. The `nanotime` requires installing another package. The first method is shown here.

In the spirit of not destroying data, we create a second lower-case column called `time`.

```{r}
# divide by 1000, convert to integer, then convert to POSIXct
data[, time := as.integer(TIME / 1000)]
```

### Add speed and turning angle

```{r}
# add incoming and outgoing speed
data[, `:=` (speed_in = atl_get_speed(data, 
                                      x = "x", 
                                      y = "y", 
                                      time = "time"),
             speed_out = atl_get_speed(data, type = "out"))]

# add turning angle
data[, angle := atl_turning_angle(data = data)]
```

### Get 95th percentile of speed and angle

```{r}
# use sapply
speed_angle_thresholds <- 
  sapply(data[, list(speed_in, speed_out, angle)], 
       quantile, probs = 0.9, na.rm = T)
```

### Plot to see speeds

```{r echo=FALSE}
# plot filtered data
fig_speed_outliers <-
  ggplot()+
  geom_point(data = data,
             aes(x, y, 
                 col = speed_in < 25.6),
             size = 0.3, alpha = 0.5,
             show.legend = F)+
  annotate(geom = "rect",
           xmin = 650785, xmax = 653250,
           ymin = 5904450, ymax = 5906133,
           fill = NA, col = "grey20")+
  annotate(geom = "rect",
           ymin = 5901650, ymax = 5903100,
           xmin = 650000, xmax = 650600,
           fill = NA, col = "grey20"
           )+
  scale_color_manual(values = c("grey", pal[3]))+
  ggthemes::theme_few()+
  theme(axis.text = element_blank(),
        axis.title = element_blank())+
  ggspatial::annotation_scale(location = "tl")+
  coord_sf(crs = 32631)+
  labs(colour = "speed (m/s)")

# save
ggsave(fig_speed_outliers, filename = "figures/fig_speed_outlier.png",
       width = 170 / 25, height = 170 / 25)
```

![](figures/fig_speed_outlier.png)

### Filter on speed

Here we use a speed threshold of 15 m/s, the fastest known boat speed.

```{r}
# make a copy
data_unproc <- copy(data)

# remove speed outliers
data <- atl_filter_covariates(data = data,
            filters = c("(speed_in < 15 & speed_out < 15)"))

# recalculate speed and angle
data[, `:=` (speed_in = atl_get_speed(data, 
                                      x = "x", 
                                      y = "y", 
                                      time = "time"),
             speed_out = atl_get_speed(data, type = "out"))]

# add turning angle
data[, angle := atl_turning_angle(data = data)]
```

## Smoothing the trajectory

```{r}
# apply a 5 point median smooth, first make a copy
data_unproc <- copy(data)

# now apply the smooth
atl_median_smooth(data = data,
                  x = "x", y = "y", time = "time",
                  moving_window = 5)
```

### Plot smoothed data

```{r}
# make zoomed in figures
fig_smooth <-
  ggplot()+
  geom_point(data = data_raw[!data_unproc, on = c("x", "y")],
             aes(x, y),
             col = "grey",
             shape = 4,
             size = 0.4)+
  # smoothed data
  geom_path(data = data,
             aes(x, y),
             col = pal[3]) +
  geom_point(data = data,
             aes(x, y),
             col = pal[3],
             shape = 19,
             alpha = 0.5) +
  coord_cartesian(
    xlim = c(650785, 653250),
    ylim = c(5904450, 5906133), 
    expand = F
  )+
  ggthemes::theme_few()+
  theme(axis.title = element_blank(),
        axis.text = element_blank())+
  ggspatial::annotation_scale(location = "br")

ggsave(fig_smooth,
       filename = "figures/fig_calib_median_smooth.png",
       width = 90 / 25, height = 90 / 25)
```

## Thinning the data

```{r}
# save a copy
data_unproc <- copy(data)

# remove columns we don't need
data <- data[, setdiff(colnames(data), 
                       c("tID", "Timestamp", "id", "TIME", "UTCtime")), 
             with = FALSE]

# thin to a 30s interval
data_thin <- atl_thin_data(data = data,
                           interval = 30,
                           method = "aggregate",
                           id_columns = "TAG")
```

### Plot thinned data

```{r echo=FALSE}
# make zoomed in figures
fig_smooth_thin <-
  ggplot()+
  geom_point(data = data_raw[!data_unproc, on = c("x", "y")],
             aes(x, y),
             col = "grey",
             shape = 4,
             size = 0.4)+

  geom_point(data = data_thin,
             aes(x, y, size = count),
             col = pal[4],
             alpha = 0.8,
             shape = 0, show.legend = F)+
  geom_point(data = data_unproc,
             aes(x, y),
             alpha = 0.6,
             col = pal[3],
             shape = 16) +
  scale_size(range = c(2, 6))+
  coord_cartesian(
    xlim = c(650785, 653250),
    ylim = c(5904450, 5906133), 
    expand = F
  )+
  ggthemes::theme_few()+
  theme(axis.title = element_blank(),
        axis.text = element_blank())+
  ggspatial::annotation_scale(location = "br")

ggsave(fig_smooth_thin,
       filename = "figures/fig_calib_smooth_thin.png",
       width = 90 / 25, height = 90 / 25, dpi = 300)
```

### Plot pre-processing steps

```{r echo=FALSE}
# make combined walkthrough figure
figure_walkthrough <-
wrap_plots(
  list(
    fig_speed_outliers, fig_smooth_thin),
  design = "A\nB")+
  plot_annotation(tag_levels = "a",
                  tag_prefix = "(",
                  tag_suffix = ")")&
  theme(plot.tag = element_text(face = "bold"))

# # save combined figure
# ggsave(figure_walkthrough,
#        filename = "figures/fig_walkthrough.png",
#        height = 170, width = 90, units = "mm")
```

![](figures/fig_walkthrough.png)

## Residence patches

### Prepare data

An indicator of individual residence at or near a position can be useful when attempting to identify residence patches. Positions can be filtered on a metric such as residence time (see Bracis et al. 2018).

### Calculate residence time

```{r}
# load recurse
library(recurse)
```

```{r}
# get 4 column data
data_for_patch <- data_thin[, list(x, y, time, TAG)]

# get recurse data for a 10m radius
recurse_stats <- getRecursions(data_for_patch,
                               radius = 50, timeunits = "mins")

# assign to recurse data
data_for_patch[, res_time := recurse_stats$residenceTime]

# save recurse data
fwrite(data_for_patch, file = "data/data_calib_for_patch.csv")
```

Subset waypoint data.

```{r}
library(stringi)
data_res <- data_unproc[stri_detect(tID, regex = "(WP)")]
```

### Run residence patch method

We subset data with a residence time > 5 minutes.

```{r}
# assign id as tag
data_for_patch[, id := as.character(TAG)]

# on known residence points
patch_res_known <- atl_res_patch(data_for_patch[res_time >= 5, ], 
                                buffer_radius = 5,
                                lim_spat_indep = 50,
                                lim_time_indep = 5,
                                min_fixes = 3)
```

### Get spatial and summary objects

```{r}
# for the known and unkniwn patches
patch_sf_data <- atl_patch_summary(patch_res_known, 
                                   which_data = "spatial",
                                   buffer_radius = 20)

# assign crs
sf::st_crs(patch_sf_data) <- 32631

# get summary data
patch_summary_data <- atl_patch_summary(patch_res_known, 
                                        which_data = "summary")
```

### Get waypoint centroids

```{r}
# get centroid
data_res_summary <- data_res[, list(x_median = median(x),
                                    y_median = median(y),
                                    t_median = median(time)),
                             by = "tID"]

# now get times 10 mins before and after
data_res_summary[, `:=`(t_min = t_median - (10 * 60),
                        t_max = t_median + (10 * 60))]

# make a list of positions 10min before and after
wp_data <- mapply(function(l, u, mx, my) {
  tmp_data <- data_unproc[inrange(time, l, u)]
  tmp_data[, distance := sqrt((mx - x)^2 + (my - y)^2)]
  # keep within 50
  tmp_data <- tmp_data[distance <= 50, ]
  # get duration
  return(diff(range(tmp_data$time)))
}, data_res_summary$t_min, data_res_summary$t_max,
   data_res_summary$x_median, data_res_summary$y_median,
SIMPLIFY = FALSE)
```

### Read Griend

```{r}
# read griend
griend <- sf::st_read("data/griend_polygon/griend_polygon.shp")
```

We prepare a plot of the residence patches.

```{r echo=FALSE}
# patch with residence points and all patches
fig_basic_residence <-
  ggplot()+
  geom_sf(data = griend, fill = "antiquewhite")+
  geom_point(data = data_res_summary, 
             aes(x_median, y_median),
             size = 10, shape = 7,
             stroke = 1,
             col = pal[4],
             alpha = 1)+
  geom_path(data = data_for_patch,
            aes(x, y),
            show.legend = F,
            size = 0.3,
            col = "grey")+
  geom_point(data = data_for_patch,
            aes(x, y),
            show.legend = F,
            shape = 1, col = "grey50")+
  geom_sf(data = patch_sf_data[patch_sf_data$duration >= 5 * 60, ],
          fill = pal[3],
          col = pal[3],
          alpha = 0.3)+
  scale_colour_manual(values = c("grey50", pal[3]))+
  scale_shape_manual(values = c(4, 19))+
  ggspatial::annotation_scale(location = "tl")+
  ggthemes::theme_few()+
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        panel.background = element_rect(fill = "aliceblue"))+
  coord_sf(crs = 32631,
           expand = F,
           ylim = c(5901650, 5903100),
           xlim = c(650000, 650600))
```

## Compare patch metrics

### Compare known patches

```{r}
# get known patch summary
data_res <- data_unproc[stringi::stri_detect(tID, regex = "(WP)"), ]

# get waypoint summary
patch_summary_real <- data_res[, list(nfixes_real = .N,
                                      x_median = round(median(x), digits = -2),
                                      y_median = round(median(y), digits = -2)), 
                               by = "tID"]

# add real duration
patch_summary_real[, duration_real := unlist(wp_data)]

# round median coordinate for inferred patches
patch_summary_inferred <- 
  patch_summary_data[, 
                     c("x_median", "y_median", 
                       "nfixes", "duration", "patch")
                     ][, `:=`(x_median = round(x_median, digits = -2),
                              y_median = round(y_median, digits = -2))]

# join with respatch summary
patch_summary_compare <- 
  merge(patch_summary_real,
        patch_summary_inferred, 
        on = c("x_median", "y_median"),
        all.x = TRUE, all.y = TRUE)

patch_summary_compare[!is.na(tID)]

# drop nas
patch_summary_compare <- na.omit(patch_summary_compare)

# drop patch around WP080
patch_summary_compare <- patch_summary_compare[tID != "WP080", ]
```

12 patches are identified where there are no waypoints, while 2 waypoints are not identified as patches. These waypoints consisted of 6 and 15 (WP098 and WP092) positions respectively, and were lost when the data were aggregated to 30 second intervals.

### Plot durations comparisons

```{r}
# get linear model
model_duration <- lm(duration_real ~ duration,
                     data = patch_summary_compare)

# get R2
summary(model_duration)

# write to file
writeLines(
  text = capture.output(
    summary(model_duration)
  ),
  con = "data/model_output_residence_patch.txt"
)
```

```{r echo=FALSE}
# make figure comparing different methods
figure_patch_duration <-
  ggplot()+
  geom_point(data = patch_summary_compare,
             aes(duration, 
                 duration_real),
             show.legend = F,
            size = 2, col = "grey")+
  geom_smooth(data = patch_summary_compare,
             aes(duration, 
                 duration_real),
             se = F,
             col = pal[1],
             lwd = 0.5,
             method = "lm",
             show.legend = F)+
  annotate(geom = "text",
           x = 600, y = 1020,
           size = 5,
           label = "R^2 == 0.95",
           parse = T, family = "Arial")+
  scale_x_continuous(
    breaks = 60 * seq(3, 18, 2),
    labels = seq(3, 18, 2)
    )+
  scale_y_continuous(
    breaks = 60 * seq(5, 20, 2),
    labels = seq(5, 20, 2)
  )+
  coord_fixed(expand = T,
              ratio = 1.115
  #             # xlim = c(400, 1050),
  #             ylim = c(250, 1050)
  )+
  geom_abline(slope = 1, lty = 2,
              col = "black")+
  ggthemes::theme_few()+
  labs(x = "inferred duration (min)",
       y = "real duration (min)")
```

## Make Figure 07

```{r echo=FALSE}
# wrap together
figure_res_patch <-
  wrap_plots(list(fig_speed_outliers,
                  fig_smooth_thin,
                  fig_basic_residence),
             design = "AC\nAC\nBC\nBC"
             )+
  plot_annotation(tag_levels = "a",
                  tag_prefix = "(",
                  tag_suffix = ")") &
  theme(plot.tag = element_text(face = "bold"))

# save figure
ggsave(figure_res_patch,
       filename = "figures/fig_07_calib_residence_patch.png",
       height = 170, width = 170, units = "mm")
```

![](figures/fig_calib_residence_patch.png)