### 2020-11-22 05_calibration_data.Rmd
make calibration supplement

---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Processing calibration data

Here we show how the residence patch method [@barraquand2008; @bijleveld2016; @oudman2019] accurately estimates the duration of known stops in a track collected as part of a calibration exercise in the Wadden Sea.

## Prepare libraries

First we prepare the libraries we need. Libraries can be installed from CRAN if necessary.

```{r}
# load libs
library(data.table)
library(atlastools)
library(ggplot2)
library(patchwork)

# prepare a palette
pal <- RColorBrewer::brewer.pal(4, "Set1")
```

## Access data and preliminary visualisation

First we access the data from a local file using the `data.table` package [@dowle2020].
We then visualise the raw data.

```{r}
# read and plot example data
data <- fread("data/atlas1060_allTrials_annotated.csv")
data_raw <- copy(data)
```

```{r echo=FALSE}
# plot data
fig_data_raw <-
  ggplot(data)+
  geom_path(aes(x, y),
             col = "grey", alpha = 1, size = 0.2)+
  geom_point(aes(x, y),
             col = "grey", alpha = 0.2, size = 0.2)+
  ggthemes::theme_few()+
  theme(axis.title = element_blank(),
        axis.text = element_blank())+
  coord_sf(crs = 32631)

# save figure
ggsave(fig_data_raw, filename = "figures/fig_calibration_raw.png",
       width = 185 / 25)
```

![](figures/fig_calibration_raw.png)

## Filter by bounding box

We first save a copy of the data, so that we can plot the raw data with the cleaned data plotted over it for comparison.

```{r}
# make a copy using the data.table copy function
data_unproc <- copy(data)
```

We then filter by a bounding box in order to remove the point outlier to the far south east of the main track. We use the `atl_filter_bounds` functions using the `x_range` argument, to which we pass the limit in the UTM 31N coordinate reference system.
This limit is used to exclude all points with an X coordinate < 645,000.

We then plot the result of filtering, with the excluded point in black, and the points that are retained in green.

```{r}
# remove inside must be set to falses
data <- atl_filter_bounds(data = data, 
                          x = "x", y = "y", 
                          x_range = c(645000, max(data$x)), 
                          remove_inside = FALSE)
```

```{r echo=FALSE}
# plot data
fig_data_bbox <-
  ggplot()+
  geom_path(data = data_raw,
            aes(x, y),
             col = "grey", alpha = 0.5, size = 0.3)+
  geom_point(data = data_raw,
             aes(x, y,
                 col = x > 645000),
             alpha = ifelse(data_raw$x > 645000, 1, 1),
             size = ifelse(data_raw$x > 645000, 0.3, 2),
             show.legend = F)+
  scale_colour_manual(values = c("black", pal[3]))+
  geom_vline(xintercept = 645000,
             col = "grey", lty = 2)+
  ggspatial::annotation_scale(location = "br")+
  ggthemes::theme_few()+
  theme(axis.text = element_blank(),
        axis.title = element_blank())+
  coord_sf(crs = 32631)

# save result
ggsave(fig_data_bbox, filename = "figures/fig_calib_bbox.png",
       width = 185 / 25)
```

## Filter trajectories

### Handle time

Time in ATLAS tracking is counted in milliseconds and is represented by a 64-bit integer (type `long`), which is not natively supported in R; it will instead be converted to a `numeric`, or `double`.

This is not what is intended, but it works. The `bit64` package can help handle 64-bit integers if you want to keep to intended type.

A further issue is that 64-bit integers (whether represented as `bit64` or `double`) do not yield meaninful results when you try to convert them to a date-time object, such as of the class `POSIXct`.

This is because `as.POSIXct` fails when trying to work with 64-bit integers (it cannot interpret this type), and returns a date many thousands of years in the future (approx. 52,000 CE) if the time column is converted to `numeric`.

There are two possible solutions. The parsimonious one is to convert the 64-bit number to a 32-bit short integer (dividing by 1000), or to use the `nanotime` package.

The conversion method loses an imperceptible amount of precision. The `nanotime` requires installing another package. The first method is shown here.

In the spirit of not destroying data, we create a second lower-case column called `time`.

```{r}
# divide by 1000, convert to integer, then convert to POSIXct
data[, time := as.integer(TIME / 1000)]
```

### Add speed and turning angle

```{r}
# add incoming and outgoing speed
data[, `:=` (speed_in = atl_get_speed(data, 
                                      x = "x", 
                                      y = "y", 
                                      time = "time"),
             speed_out = atl_get_speed(data, type = "out"))]

# add turning angle
data[, angle := atl_turning_angle(data = data)]
```

### Get 95th percentile of speed and angle

```{r}
# use sapply
speed_angle_thresholds <- 
  sapply(data[, list(speed_in, speed_out, angle)], 
       quantile, probs = 0.9, na.rm = T)
```

```{r echo=FALSE}
# plot filtered data
fig_speed_outliers <-
  ggplot()+
  geom_point(data = data,
             aes(x, y, 
                 col = speed_in < 25.6),
             size = 0.3, alpha = 0.5,
             show.legend = F)+
  annotate(geom = "rect",
           xmin = 650785, xmax = 653250,
           ymin = 5904450, ymax = 5906133,
           fill = NA, col = "grey20")+
  annotate(geom = "rect",
           ymin = 5901650, ymax = 5903100,
           xmin = 650000, xmax = 650600,
           fill = NA, col = "grey20"
           )+
  scale_color_manual(values = c("grey", pal[3]))+
  ggthemes::theme_few()+
  theme(axis.text = element_blank(),
        axis.title = element_blank())+
  ggspatial::annotation_scale(location = "tl")+
  coord_sf(crs = 32631)+
  labs(colour = "speed (m/s)")

# save
ggsave(fig_speed_outliers, filename = "figures/fig_speed_outlier.png",
       width = 170 / 25, height = 170 / 25)
```

### Filter on speed

Here we use a speed threshold of 15 m/s, the fastest known boat speed.
We then plot the data with the extreme speeds shown in grey, and the positions retained shown in green.

```{r}
# make a copy
data_unproc <- copy(data)

# remove speed outliers
data <- atl_filter_covariates(data = data,
            filters = c("(speed_in < 15 & speed_out < 15)"))

# recalculate speed and angle
data[, `:=` (speed_in = atl_get_speed(data, 
                                      x = "x", 
                                      y = "y", 
                                      time = "time"),
             speed_out = atl_get_speed(data, type = "out"))]

# add turning angle
data[, angle := atl_turning_angle(data = data)]
```

![](figures/fig_speed_outlier.png)

## Smoothing the trajectory

We then apply a median smooth over a moving window ($K$ = 5).
This function modifies in place, and does not need to be assigned to a new variable.
We create a copy of the data before applying the smooth so that we can compare the data before and after smoothin.

```{r}
# apply a 5 point median smooth, first make a copy
data_unproc <- copy(data)

# now apply the smooth
atl_median_smooth(data = data,
                  x = "x", y = "y", time = "time",
                  moving_window = 5)
```

```{r}
# make zoomed in figures
fig_smooth <-
  ggplot()+
  geom_point(data = data_raw[!data_unproc, on = c("x", "y")],
             aes(x, y),
             col = "grey",
             shape = 4,
             size = 0.4)+
  geom_point(data = data,
             aes(x, y),
             col = pal[3],
             shape = 19,
             alpha = 0.5) +
  coord_cartesian(
    xlim = c(650785, 653250),
    ylim = c(5904450, 5906133), 
    expand = F
  )+
  ggthemes::theme_few()+
  theme(axis.title = element_blank(),
        axis.text = element_blank())+
  ggspatial::annotation_scale(location = "br")

ggsave(fig_smooth,
       filename = "figures/fig_calib_median_smooth.png",
       width = 90 / 25, height = 90 / 25)
```

![](figures/fig_calib_median_smooth.png)

## Thinning the data

Next we thin the data to demonstrate thinning by median smoothing.
Following this, we plot the median smooth and thinning by aggregation.

```{r}
# save a copy
data_unproc <- copy(data)

# remove columns we don't need
data <- data[, setdiff(colnames(data), 
                       c("tID", "Timestamp", "id", "TIME", "UTCtime")), 
             with = FALSE]

# thin to a 30s interval
data_thin <- atl_thin_data(data = data,
                           interval = 30,
                           method = "aggregate",
                           id_columns = "TAG")
```

```{r echo=FALSE}
# make zoomed in figures
fig_smooth_thin <-
  ggplot()+
  geom_point(data = data_raw[!data_unproc, on = c("x", "y")],
             aes(x, y),
             col = "grey",
             shape = 4,
             size = 0.4)+

  geom_point(data = data_thin,
             aes(x, y, size = count),
             col = pal[4],
             alpha = 0.8,
             shape = 0, show.legend = F)+
  geom_point(data = data_unproc,
             aes(x, y),
             alpha = 0.6,
             col = pal[3],
             shape = 16) +
  scale_size(range = c(2, 6))+
  coord_cartesian(
    xlim = c(650785, 653250),
    ylim = c(5904450, 5906133), 
    expand = F
  )+
  ggthemes::theme_few()+
  theme(axis.title = element_blank(),
        axis.text = element_blank())+
  ggspatial::annotation_scale(location = "br")

ggsave(fig_smooth_thin,
       filename = "figures/fig_calib_smooth_thin.png",
       width = 90 / 25, height = 90 / 25, dpi = 300)
```

```{r echo=FALSE}
# make combined walkthrough figure
figure_walkthrough <-
wrap_plots(
  list(
    fig_speed_outliers, fig_smooth_thin),
  design = "AB")+
  plot_annotation(tag_levels = "a",
                  tag_prefix = "(",
                  tag_suffix = ")")&
  theme(plot.tag = element_text(face = "bold"))

# # save combined figure
ggsave(figure_walkthrough,
       filename = "figures/fig_walkthrough.png",
       height = 90, width = 170, units = "mm")
```

![](figures/fig_calib_smooth_thin.png)

## Residence patches

### Get waypoint centroids

We subset the annotated calibration data to select the waypoints and the positions around them which are supposed to be the locations of known stops. Since each stop was supposed to be 5 minutes long, there are multiple points in each known stop.

```{r}
library(stringi)
data_res <- data_unproc[stri_detect(tID, regex = "(WP)")]
```

From this data, we get the centroid of known stops, and determine the time difference between the first and last point within 50 metres, and within 10 minutes of the waypoint positions' median time.

Essentially, this means that the maximum duration of a stop can be 20 minutes, and stops above this duration are not expected.

```{r}
# get centroid
data_res_summary <- data_res[, list(x_median = round(median(x), digits = -2),
                                    y_median = round(median(y), digits = -2)),
                                    t_median = median(time)),
                             by = "tID"]

# now get times 10 mins before and after
data_res_summary[, `:=`(t_min = t_median - (10 * 60),
                        t_max = t_median + (10 * 60))]

# make a list of positions 10min before and after
wp_data <- mapply(function(l, u, mx, my) {
  tmp_data <- data_unproc[inrange(time, l, u)]
  tmp_data[, distance := sqrt((mx - x)^2 + (my - y)^2)]

  # keep within 50
  tmp_data <- tmp_data[distance <= 50, ]

  # get duration
  return(diff(range(tmp_data$time)))

}, data_res_summary$t_min, data_res_summary$t_max,
   data_res_summary$x_median, data_res_summary$y_median,
SIMPLIFY = TRUE)
```

### Prepare data

An indicator of individual residence at or near a position can be useful when attempting to identify residence patches. Positions can be filtered on a metric such as residence time [@bracis2018].

### Calculate residence time

First we calculate the residence time with a radius of 50 metres.
For this, we need a dataframe with coordinates, the timestamp, and the animal id.
We save this data to file for later use.

```{r}
# load recurse
library(recurse)
```

```{r}
# get 4 column data
data_for_patch <- data_thin[, list(x, y, time, TAG)]

# get recurse data for a 10m radius
recurse_stats <- getRecursions(data_for_patch,
                               radius = 50, timeunits = "mins")

# assign to recurse data
data_for_patch[, res_time := recurse_stats$residenceTime]

# save recurse data
fwrite(data_for_patch, file = "data/data_calib_for_patch.csv")
```

### Run residence patch method

We subset data with a residence time > 5 minutes in order to construct residence patches.
From this subset, we construct residence patches using the parameters: `buffer_radius` = 5 metres, `lim_spat_indep` = 50 metres, `lim_time_indep` = 5 minutes, and `min_fixes` = 3.

```{r}
# assign id as tag
data_for_patch[, id := as.character(TAG)]

# on known residence points
patch_res_known <- atl_res_patch(data_for_patch[res_time >= 5, ], 
                                buffer_radius = 5,
                                lim_spat_indep = 50,
                                lim_time_indep = 5,
                                min_fixes = 3)
```

### Get spatial and summary objects

We get spatial and summary ouput of the residence patch method using the `atl_patch_summary` function using the options `which_data` = "spatial" and `which_data` = "summary.
We use a buffer radius here of 20 metres for the spatial buffer, despite using a buffer radius of 5 metres earlier, simply because it is easier to visualise in the output figure.

```{r}
# for the known and unkniwn patches
patch_sf_data <- atl_patch_summary(patch_res_known, 
                                   which_data = "spatial",
                                   buffer_radius = 20)

# assign crs
sf::st_crs(patch_sf_data) <- 32631

# get summary data
patch_summary_data <- atl_patch_summary(patch_res_known, 
                                        which_data = "summary")
```

### Prepare to plot data

We read in the island's shapefile to plot it as a background for the residence patch figure.

```{r}
# read griend
griend <- sf::st_read("data/griend_polygon/griend_polygon.shp")
```

```{r echo=FALSE}
# patch with residence points and all patches
fig_basic_residence <-
  ggplot()+
  geom_sf(data = griend, fill = "antiquewhite")+
  geom_point(data = data_res_summary, 
             aes(x_median, y_median),
             size = 10, shape = 7,
             stroke = 1,
             col = pal[4],
             alpha = 1)+
  geom_path(data = data_for_patch,
            aes(x, y),
            show.legend = F,
            size = 0.3,
            col = "grey")+
  geom_point(data = data_for_patch,
            aes(x, y),
            show.legend = F,
            shape = 1, col = "grey50")+
  geom_sf(data = patch_sf_data[patch_sf_data$duration >= 5 * 60, ],
          fill = pal[3],
          col = pal[3],
          alpha = 0.3)+
  scale_colour_manual(values = c("grey50", pal[3]))+
  scale_shape_manual(values = c(4, 19))+
  ggspatial::annotation_scale(location = "tl")+
  ggthemes::theme_few()+
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        panel.background = element_rect(fill = "aliceblue"))+
  coord_sf(crs = 32631,
           expand = F,
           ylim = c(5901650, 5903100),
           xlim = c(650000, 650600))
```

## Compare patch metrics

We then merge the annoated, known stop data with the calculated patch duration.
We filter this data to exclude one exceedingly long outlier of about an hour (WP080), which how

```{r}
# get known patch summary
data_res <- data_unproc[stringi::stri_detect(tID, regex = "(WP)"), ]

# get waypoint summary
patch_summary_real <- data_res[, list(nfixes_real = .N,
                                      x_median = round(median(x), digits = -2),
                                      y_median = round(median(y), digits = -2)), 
                               by = "tID"]

# add real duration
patch_summary_real[, duration_real := wp_data]

# round median coordinate for inferred patches
patch_summary_inferred <- 
  patch_summary_data[, 
                     c("x_median", "y_median", 
                       "nfixes", "duration", "patch")
                     ][, `:=`(x_median = round(x_median, digits = -2),
                              y_median = round(y_median, digits = -2))]

# join with respatch summary
patch_summary_compare <- 
  merge(patch_summary_real,
        patch_summary_inferred, 
        on = c("x_median", "y_median"),
        all.x = TRUE, all.y = TRUE)

# drop nas
patch_summary_compare <- na.omit(patch_summary_compare)

# drop patch around WP080
patch_summary_compare <- patch_summary_compare[tID != "WP080", ]
```

7 patches are identified where there are no waypoints, while 2 waypoints are not identified as patches. These waypoints consisted of 6 and 15 (WP098 and WP092) positions respectively, and were lost when the data were aggregated to 30 second intervals.

### Linear model durations

We run a simple linear model.

```{r}
# get linear model
model_duration <- lm(duration_real ~ duration,
                     data = patch_summary_compare)

# get R2
summary(model_duration)

# write to file
writeLines(
  text = capture.output(
    summary(model_duration)
  ),
  con = "data/model_output_residence_patch.txt"
)
```

```{r echo=FALSE}
# make figure comparing different methods
figure_lm_duration <-
  ggplot()+
  geom_point(data = patch_summary_compare,
             aes(duration, 
                 duration_real),
             show.legend = F,
            size = 2, col = "grey")+
  geom_smooth(data = patch_summary_compare,
             aes(duration, 
                 duration_real),
             se = F,
             col = pal[1],
             lwd = 0.5,
             method = "lm",
             show.legend = F)+
  annotate(geom = "text",
           x = 600, y = 1020,
           size = 5,
           label = "R^2 == 0.908",
           parse = T, family = "Arial")+
  scale_x_continuous(
    breaks = 60 * seq(3, 18, 2),
    labels = seq(3, 18, 2)
    )+
  scale_y_continuous(
    breaks = 60 * seq(5, 20, 2),
    labels = seq(5, 20, 2)
  )+
  coord_fixed(expand = T,
              ratio = 1.115
  #             # xlim = c(400, 1050),
  #             ylim = c(250, 1050)
  )+
  geom_abline(slope = 1, lty = 2,
              col = "black")+
  ggthemes::theme_few()+
  labs(x = "inferred duration (min)",
       y = "real duration (min)")

ggsave(figure_lm_duration, filename = "figures/fig_calib_lm_duration.png",
       width = 80 / 25, height = 80 / 25)
```

![](figures/fig_calib_lm_duration.png)

```{r echo=FALSE}
# wrap together
figure_res_patch <-
  wrap_plots(list(fig_speed_outliers,
                  fig_smooth_thin,
                  fig_basic_residence),
             design = "AC\nAC\nBC\nBC"
             )+
  plot_annotation(tag_levels = "a",
                  tag_prefix = "(",
                  tag_suffix = ")") &
  theme(plot.tag = element_text(face = "bold"))

# save figure
ggsave(figure_res_patch,
       filename = "figures/fig_07_calib_residence_patch.png",
       height = 170, width = 170, units = "mm")
```