### 2021-10-07 supplement/index.Rmd
add index

--- 
knit: "bookdown::render_book"
title: "Supplementary Material for _A Guide to Pre-processing High-Frequency Animal Tracking Data_"
author:
    - Pratik R. Gupte
    - Christine E. Beardsworth
    - Orr Spiegel
    - Emmanuel Lourie
    - Sivan Toledo
    - Ran Nathan
    - Allert I. Bijleveld
date: "`r Sys.Date()`"
output: 
    bookdown::pdf_document2
    bookdown::gitbook
documentclass: scrreprt
toc-depth: 1
secnumdepth: 2
geometry: "left=4cm, right=3cm, top=2.5cm, bottom=2.5cm"
bibliography: [references.bib]
link-citations: yes
github-repo: pratikunterwegs/atlas-best-practices
description: "Supplementary Material for _A Guide to Pre-processing High-Frequency Animal Tracking Data_"
---

# Introduction {-}

This is a rendering of the Supplementary Material for _A Guide to Pre-processing High-Frequency Animal Tracking Data_, and contains two fully worked out examples that could provide a useful template for structuring pre-processing pipelines for your own high-throughput tracking data.

The code and data used here are available on Github at [github.com/pratikunterwegs/atlas-best-practices](https://github.com/pratikunterwegs/atlas-best-practices), and on Zendo at [zenodo/atlas-best-practices](https://doi.org/10.5281/zenodo.4287462).